---
title: "Homework 2 Cereal Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 1. Read in data: 

```{r}
#LOAD NECESSARY PACKAGES
library(R2jags)
library(lattice)
library(knitr)
library(ggplot2)

setwd("/Users/huiyuhu/Desktop/Study/UCLA_Biostat/BIOSTAT234/HW/HW2")
getwd()
load("Cereal_Data.RData")

# take look at data
table1 <- head(Cereal_Data)
kable(table1)

is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}
library(dplyr)
Cereal_Data %>%
  mutate(outlier = ifelse(is_outlier(unitcal), unitcal, as.numeric(NA))) %>%
  ggplot(., aes(x = "", y = unitcal)) +
    geom_boxplot() +
    geom_text(aes(label = ), na.rm = TRUE, hjust = -0.3)

Cereal_Data$unitcal <- Cereal_Data$calories/Cereal_Data$`size(cups)`
ggplot(data = Cereal_Data, aes(x = "", y = unitcal, label = rownames(Cereal_Data))) + 
  geom_boxplot() + geom_text() + labs(title="Boxplot of Unit Calories")

Cereal_Data <- Cereal_Data[-44, ]

```
- Variables in the data set are the cost of the box of cereal in dollars, name of the cereal, the weight (ounces) of the cereal, the calories per serving, the serving size (cups), the manufacturer's name and a coding of name into the integers 1 to 8 and finally whether the cereal box has a cartoon character on the front (yes = 1, no = 0). 

- The goal is to predict cost. In particular, do caloric content, adult/kid target audience and manufacturer matter for price?
```{r}
1/var(Cereal_Data$`cost(dollars)`)
```
```{r}

```

#### 2. Fit the model

```{r}
load("AddBurnin.RData")
library(lattice)
load("Cereal_Data.RData")
data <- Cereal_Data[order(Cereal_Data$company_numbers), ]
table <- data.frame(table(data$company_numbers))[, 2]

y <- data$`cost(dollars)`
x <- data[, c("weight(ounces)", "unitcal", "size(cups)", "cartooncharacter")]
x <- cbind(rep(1, 46), x)
N <- 46

#Create the model
sink("hw2model.txt")
cat("
model
        {       
                for( i in 1 : 46 ) {
                            y[i] ~ dnorm(mu[i], tau.e)
                                mu[i] <- inprod(x[i, ], alpha[]) + beta[comp[i]]
                }

                for (j in 1 : 8) {
                       beta[j] ~ dnorm(0, tau.b)
                }

for( k in 1:5) {
                alpha[k]~dnorm(0, 0.001)
}

                tau.e ~ dgamma(ea,eb)
                tau.b ~ dgamma(ba,bb)

                SE_e <- 1 /sqrt(tau.e)
                SE_b <- 1 /sqrt(tau.b)
                rho <- SE_b*SE_b/(SE_e*SE_e + SE_b*SE_b)


        }

    ",fill = TRUE)
sink()

sink("hw2model.t.txt")
cat("
model
        {       
                for( i in 1 : 46 ) {
                            rbrand[j] ~ dt(alpha, tau.b,df2)
                            cost[i] ~ dt(mu[i],tau.e,df1)
                            df1 <- 1/invdf1
                            invdf1 ~ dunif(0,.5)
                            df2 <- 1/invdf2
                            invdf2 ~ dunif(0,.5)

                }

                for (j in 1 : 8) {
                       beta[j] ~ dnorm(0, tau.b)
                }

for( k in 1:5) {
                alpha[k]~dnorm(0, 0.001)
}

                tau.e ~ dgamma(ea,eb)
                tau.b ~ dgamma(ba,bb)

                SE_e <- 1 /sqrt(tau.e)
                SE_b <- 1 /sqrt(tau.b)
                rho <- SE_b*SE_b/(SE_e*SE_e + SE_b*SE_b)


        }

    ",fill = TRUE)
sink()


##inits setup:
inits <- rep(list(list(
  tau.e = 1,
  tau.b = 1,
  beta = rep(0, 8),
  alpha = c(1, 1, 1, 1, 1)
)), 5)

#inits <- rep(list(list(yinits = rep(0, length(y)), tau.e, tau.b, beta, alpha)), 5)
##priordata:
priordata = list(
  comp  = data$company_numbers,
  ea     = 2, 
  eb     = 1.86, 
  ba     = 5, 
  bb     = 2.5,
  x      = x,
  y	     = y
)
parameters = c(
  "alpha", "beta", "tau.e", 
  "tau.b", "SE_e", "SE_b")
proc.time()
run1 = jags(data = priordata, inits = inits, parameters.to.save = parameters, "hw2model.txt", 
	n.chains=5, n.iter=11000, n.burnin=0, n.thin=1)
proc.time()
# 1100 iterations takes about 3 seconds on my computer.  
# 11000 iterations takes a little under 15 seconds on my computer.  

names(run1)
Output1=AddBurnin(run1$BUGSoutput$sims.array,burnin=1000,n.thin=1)

print(Output1$Burnin.Summary)


```
```{r}
par(mfrow=c(2,3))
acf(run1$BUGSoutput$sims.array[1001:11000,1,3], main="intercept", lag.max = 50)
acf(run1$BUGSoutput$sims.array[1001:11000,1,4], main="weight", lag.max = 50)
acf(run1$BUGSoutput$sims.array[1001:11000,1,5], main="calories", lag.max = 50)
acf(run1$BUGSoutput$sims.array[1001:11000,1,6], main="size", lag.max = 50)
acf(run1$BUGSoutput$sims.array[1001:11000,1,7], main="cartoon", lag.max = 50)
par(mfrow=c(3,3))
acf(run1$BUGSoutput$sims.array[1001:11000,1,8], main="beta1", lag.max = 50)
acf(run1$BUGSoutput$sims.array[1001:11000,1,9], main="beta2", lag.max = 50)
acf(run1$BUGSoutput$sims.array[1001:11000,1,10], main="beta3", lag.max = 50)
acf(run1$BUGSoutput$sims.array[1001:11000,1,11], main="beta4", lag.max = 50)
acf(run1$BUGSoutput$sims.array[1001:11000,1,12], main="beta5", lag.max = 50)
acf(run1$BUGSoutput$sims.array[1001:11000,1,13], main="beta6", lag.max = 50)
acf(run1$BUGSoutput$sims.array[1001:11000,1,14], main="beta7", lag.max = 50)
acf(run1$BUGSoutput$sims.array[1001:11000,1,15], main="beta8", lag.max = 50)

# print run1 to get DIC = 131.8

# trace plot 
par(mfrow=c(2,3))
plot(run1$BUGSoutput$sims.array[1001:11000,1,3], main="intercept", type="l",ylab = "")
plot(run1$BUGSoutput$sims.array[1001:11000,1,4], main="weight", type="l",ylab = "")
plot(run1$BUGSoutput$sims.array[1001:11000,1,5], main="calories", type="l",ylab = "")
plot(run1$BUGSoutput$sims.array[1001:11000,1,6], main="size", type="l",ylab = "")
plot(run1$BUGSoutput$sims.array[1001:11000,1,7], main="cartoon", type="l",ylab = "")
plot(run1$BUGSoutput$sims.array[1001:11000,1,8], main="beta1", type="l",ylab = "")

plot(run1$BUGSoutput$sims.array[1001:11000,1,9], main="beta2", type="l",ylab = "")
plot(run1$BUGSoutput$sims.array[1001:11000,1,10], main="beta3", type="l",ylab = "")
plot(run1$BUGSoutput$sims.array[1001:11000,1,11], main="beta4", type="l",ylab = "")
plot(run1$BUGSoutput$sims.array[1001:11000,1,12], main="beta5", type="l",ylab = "")
plot(run1$BUGSoutput$sims.array[1001:11000,1,13], main="beta6", type="l",ylab = "")

plot(run1$BUGSoutput$sims.array[1001:11000,1,14], main="beta7", type="l",ylab = "")
plot(run1$BUGSoutput$sims.array[1001:11000,1,15], main="beta8", type="l",ylab = "")



```
```{r}
#Treat the first 1000 iterations as a burn in   
Output.hw2 = AddBurnin(run1$BUGSoutput$sims.array,burnin=1000,n.thin=2)

names1 <- paste0("beta[", 1:8, "]")
names2 <- paste0("alpha[", 1:5, "]")
Parameter2 <- c("intercept", "weight", "calories", "size", "cartoon")
Parameter <- c("365 everyday value", "Cascadian Farm", "General Mills", "kashi", "kelloggs", "Nature's Path", "post", "quaker")
table1 <- cbind(Parameter, round(Output.hw2$Burnin.Summary[names1, ], 4))
table2 <- cbind(Parameter2, round(Output.hw2$Burnin.Summary[names2, ], 4))
knitr::kable(table1, caption = "Posterior of Betas (Brand)")
knitr::kable(table2, caption = "Posterior of Alphas")
#Boxplots for Beta ## should include in the hw2
boxplot(Output.hw2$Burnin.sims.matrix[,c(seq(8,15))], range=0, axes=F)
axis(side=1,at=seq(1,62,6), labels=paste("Beta",seq(1,62,6)))
axis(side=2, at=c(-5,0,5), labels=c(-5,0,5))
box()
```

