---
title: "Lab 4"
output: html_document
---
##### Name: Huiyu Hu

```{r}
#CHANGE WORKING DIRECTORY
setwd("/Users/huiyuhu/Desktop/Study/UCLA_Biostat/BIOSTAT234/lab/Lab 4")
getwd()

#LOAD NECESSARY PACKAGES
library(R2jags)
load("AddBurnin.RData")
library(lattice)
library(knitr)
# useful function
mysummary = function(invector) {
c(mean(invector), sd(invector), quantile(invector, .025), 
	quantile(invector,.975),
	length(invector[invector>0])/length(invector))
}

# load the data.  

load("lab4_data.rdata")
head(lab4_data)
```
```{r}
#Create the model

# sink("lab4model.txt")
# cat("
# model
#         {               
#                 for( i in 1 : 64 ) {
#                         for( j in 1 : 4 ) {
#                             s[i, j]<-4*(i-1)+j
#                             y[i, j] ~ dnorm(mu[i , j],tau.e)
#                                 mu[i , j] <- inprod(x[s[i,j],],alpha[])+beta[i]
#                         }
#                         beta[i]~dnorm(0, tau.b)
#                 }
# 
# for( k in 1:8) {
#                 alpha[k]~dnorm(m[k],varinv[k])
#                 alphasign[k] <- step(alpha[k])
# }
# 
#                 tau.e ~ dgamma(ea,eb)
#                 tau.b~dgamma(ba,bb)
# 
#                 sigma <- 1 /sqrt( tau.e)
#                 sqrtD <- 1 /sqrt( tau.b)
#                 rho <- sqrtD*sqrtD/(sigma*sigma + sqrtD *sqrtD)
# 
#         }
# 
#     ",fill = TRUE)
# sink()
```

```{r}
run1 = jags(priordata, inits, parameters, "lab4model.txt", 
	n.chains=5, n.iter=1100, n.burnin=0, n.thin=1)
names(run1)
Output1=AddBurnin(run1$BUGSoutput$sims.array,burnin=100,n.thin=1)
print(Output1$Burnin.Summary)
```
```{r}
# sink("lab4Tmodel.txt")
# cat("
# model
#         {               
#                 for( i in 1 : 64 ) {
#                         for( j in 1 : 4 ) {
#                             s[i, j]<-4*(i-1)+j
#                             y[i, j] ~ dt(mu[i , j],tau.e, df1) # change to t
# 
#                             mu[i , j] <- inprod(x[s[i,j],],alpha[])+beta[i]
#                         }
#                         beta[i]~dt(0, tau.b, df2) # change to t 
# 
#                 }
# 
# for( k in 1:8) {
#                 alpha[k]~dnorm(m[k],varinv[k])
#                 alphasign[k] <- step(alpha[k])
# }
# 
#                 df1 <- 1/invdf1
#                 invdf1 ~ dunif(0, 0.5)
#                 df2 <- 1/invdf2
#                 invdf2 ~ dunif(0, 0.5)
#             
#                 tau.e ~ dgamma(ea,eb)
#                 tau.b~dgamma(ba,bb)
# 
#                 sigma <- 1 /sqrt( tau.e)
#                 sqrtD <- 1 /sqrt( tau.b)
#                 rho <- sqrtD*sqrtD/(sigma*sigma + sqrtD *sqrtD)
#         }
# 
#     ",fill = TRUE)
# 
# sink()
```
```{r}
# T
parameters = c("alpha", "alphasign", "tau.e", "tau.b", "sigma", "sqrtD", "rho", "beta[1:5]", "y[4,3:4]", "df1", "df2") 
run2 = jags(priordata, inits, parameters, "lab4Tmodel.txt",  n.chains=5, n.iter=11000, n.burnin=0, n.thin=1)
Output2 = AddBurnin(run2$BUGSoutput$sims.array,burnin=1000,n.thin=1)
```


1.	Turn in results from the t-model.  Be sure to run sufficient iterations.  
a.	How is the convergence?  Show an illustrative autocorrelation function and time-series plot for two parameters of interest.  
```{r}
temp2= Output2$Burnin.sims.array
# par(mfrow = c(1,2))
# acf(temp2[,1,"alpha[2]"], main="") 
# acf(temp2[1:5000,1,1], main="beta1", lag.max = 200)

# auto-correlation 
par(mfrow=c(2,2))
acf(temp2[1:5000,1,3], main="Auto Correlation of Interaction AA", lag.max = 30)
acf(temp2[1:5000,1,4], main="Interaction AD", lag.max = 30)

plot(temp2[1:1000,1,3], main="Trace of Interaction AA", type ="l")
plot(temp2[1:1000,1,4], main="Interaction AD", type ="l")



```

* The parameters shown above converge well.

b.	Turn in a table of results for the fixed effects, the two standard deviations sqrtD and ï³, and the two degrees of freedom parameters.  Label rows appropriately and format the table carefully.  
```{r}
temp2 = Output2$Burnin.Summary
temp = Output1$Burnin.Summary
desired_rows = c(paste0("alpha[", 1:8, "]"), "sigma", "sqrtD", "df1", "df2")
table = round(temp2[desired_rows, ],3)
table2 = temp[c(paste0("alpha[", 1:8, "]"), "sigma", "sqrtD"), ]
colnames(table2) = c("mean", "sd","2.5%", "97.5%", "P>0")
kable(table)
```
2. Compare the results from the normal model to the results from the t model:  What changes are there?  In particular, what scientific conclusions change?  

* According to the result, most of the parameter do not have big difference, expect alpha[7] (Normal: 0.39 -> t-model: 0.204). Therefore, the benefit of teaching distractors to distract may be less significant than estimated by the normal model. 


3. Reproduce figures 1-5 (see below for the normal model figures) for your t model. Label your figures appropriately. 

```{r}
out.q3 = Output2$Burnin.sims.matrix
# Plot 1
plot(density((out.q3[,"y[4,3]"])), xlim = c(0, 8), ylim = c(0, 1.5), col = "blue", lty = 2, xlab = "ln(time) (seconds)", main = "")
lines(density((out.q3[,"y[4,4]"])), col = "green", lty = 1)
legend("topright", col = c("blue","green"), legend = c('y[4,3]','y[4,4]'), bty = 'n', lty = c(2,1))
title("1. Predictions")
```

```{r}
# Plot 2
plot(density(exp(out.q3[,c("alpha[1]")] + 
               out.q3[,c("alpha[2]")])), ylim = c(0, .17), 
     main = "2. Baseline", col = "green", lty = 2)
lines(density(exp(out.q3[,c("alpha[1]")])), col = "blue", lty = 1)
legend("topright", col = c("blue","green"), legend = c('Attender','Distractor'), bty = 'n', lty = c(1,2), lwd = 2)
```
```{r}
# Plot 3
# Treatment effect for attenders
# Log scale
par(mfrow = c(1,2), oma = c(0,0,4,0), mar = c(5,3,1,0))
plot(density((out.q3[, "alpha[3]"])), main = "Effects on attenders", lty = 1)
lines(density((out.q3[,"alpha[4]"])), col = "red", lty = 2)
lines(density((out.q3[,"alpha[5]"])), col = "blue", lty = 4)
plot(density((out.q3[, "alpha[6]"])), main = "Effects on distracters", lty = 1, xlim = c(-1, 1.5))
lines(density((out.q3[,"alpha[7]"])), col = "red", lty = 2)
lines(density((out.q3[,"alpha[8]"])), col = "blue", lty = 4)
legend("topright", col = c("black", "red", "blue"), legend = c("attend", "distract", "null"), lty = c(1,2,4), bty = "n", cex = 0.8)
title("3. Treatment effect ", outer = T)
```
```{r}
# Plot 4
# Multiplicative scale
par(mfrow = c(1,2), oma = c(0,0,4,0), mar = c(5,3,1,0))
plot(density(exp(out.q3[, "alpha[3]"])), main = "Effects on attenders", xlab = "seconds", ylim = c(0, 4), lty = 1)
lines(density(exp(out.q3[,"alpha[4]"])), col = "red", lty = 2)
lines(density(exp(out.q3[,"alpha[5]"])), col = "blue", lty = 4)
plot(density(exp(out.q3[, "alpha[6]"])), main = "Effects on distracters", xlab = "seconds", xlim = c(0,2.5), lty = 1)
lines(density(exp(out.q3[,"alpha[7]"])), col = "red", lty = 2)
lines(density(exp(out.q3[,"alpha[8]"])), col = "blue", lty = 4)
title("4. Treatment effect by personality", outer = T)
legend("topright", col = c("black", "red", "blue"), legend = c("attend", "distract", "null"), lty = c(1,2,4), bty = "n", cex = 0.8)
```

```{r}
# Plot 5
# Difference
par(mfrow = c(1,2), oma = c(0,0,4,0), mar = c(5,3,1,0))
plot(density((out.q3[, "alpha[4]"]-out.q3[,"alpha[3]"])), main = "Differences for attenders", xlab = "attenders", ylim = c(0, 3), lty = 1)
lines(density((out.q3[,"alpha[4]"]-out.q3[,"alpha[5]"])), col = "red", lty = 2)
lines(density((out.q3[,"alpha[3]"]-out.q3[,"alpha[5]"])), col = "blue", lty = 4)
legend("topleft", col = c("black", "red", "blue"), legend = c("D-A", "D-N", "A-N"), lty = c(1,2,4), bty = "n", cex = 0.8)
plot(density((out.q3[, "alpha[7]"]-out.q3[,"alpha[6]"])), main = "Differences for distracter", xlab = "distracter", ylim = c(0, 2.5), xlim = c(-1.5, 1.75), lty = 1)
lines(density((out.q3[,"alpha[7]"]-out.q3[,"alpha[8]"])), col = "red", lty = 2)
lines(density((out.q3[,"alpha[6]"]-out.q3[,"alpha[8]"])), col = "blue", lty = 4)
legend("topleft", col = c("black", "red", "blue"), legend = c("D-A", "D-N", "A-N"), lty = c(1,2,4), bty = "n", cex = 0.8)
title("5. Differences in treatment effect" ,outer = T)
```

4. Invent another prior for the df, and in one sentence explain its properties (ie support, mean, sd or other characteristics) and why it is better than the above prior. 

* I suggust that df = Gamma(3969/200, 63/200), since the distribution has mean of 63 (df = n-1 for y in our dataset), also has sd of 200. It may become better since the mean will become closer to sample size and keep wide range of df.

