---
title: "Lab 4.5"
output: html_document
---
```{r}
setwd("/Users/huiyuhu/Desktop/Study/UCLA_Biostat/BIOSTAT234/lab/Lab 4_5")
getwd()
expit = function(a) exp(a)/(1+exp(a))
logit = function(a) log(a/(1-a))
#LOAD NECESSARY PACKAGES
library(R2jags)
library(knitr)
load("AddBurnin.RData")
library(lattice)
#install.packages("bayesplot")
library("bayesplot")
#install.packages("ggmcmc")
library("ggmcmc")
library("ggplot2")
library(xtable)
# useful function
mysummary = function(invector) {
  c(mean(invector), sd(invector), quantile(invector, .025), 
    quantile(invector,.975),
    length(invector[invector>0])/length(invector))
}
```

* Data

```{r}
metadata = matrix(data = c(
    506,     21,    164,     20,
  20614,    222,   1575,     59,
  32279,    681,   3051,    212,
    234,     49,     59,     19,
    201,     28,     30,     11,
    568,     60,    116,     19,
   2035,    130,    549,     43,
    406,     32,     99,     17,
   2946,     55,    831,     83), byrow=T, ncol=4)
#metadata
colnames(metadata) = c("n0.HD", "y0.HD", "n1.HD", "y1.HD")
dim(metadata)
kable(metadata) # table output in latex format
```
* brief exploration of the data

```{r}
zzz = cbind(metadata[,2]/metadata[,1],metadata[,4]/metadata[,3])
ORs = (zzz[,2]/(1-zzz[,2]))/ (zzz[,1]/(1-zzz[,1]))
zzz = cbind(zzz, ORs)
zzz = round(zzz, 3)
kable(zzz)

```
```{r}
# #Meta-analysis model 1
# sink("syncope1.txt")
# cat("
#     model
#     {
#     for( i in 1:npapers ) {
#       y0[i] ~ dbin(pie0[i],n0[i])
#       y1[i] ~ dbin(pie1[i],n1[i])
#       logit(pie0[i]) = alpha + beta[i] - delta[i]/2
#       logit(pie1[i]) = alpha + beta[i] + delta[i]/2
#       beta[i]  ~ dnorm(0 , sigmainv2)
#       delta[i] ~ dnorm(d0, tauinv2  )
#       OR[i]    = (pie1[i]/(1 - pie1[i])) / (pie0[i]/(1 - pie0[i]))
#     }
#     alpha ~ dnorm(a, b)
#     d0    ~ dnorm(0, d)
#     sigmainv2 ~ dgamma(c1,c2)
#     tauinv2   ~ dgamma(f1,f2)
#   sigma = 1/sqrt(sigmainv2)
#   tau   = 1/sqrt(tauinv2) 
#     }
#     ",fill = TRUE)
# sink()
```

```{r}
# Prior parameters
npapers = 9
a = -2.75  #
b = 1/2
d = 1/2
c1 = f1 = 3
c2 = f2 = 2


parameters= c(
  "pie0", "pie1", "alpha", "sigma", 
  "tau", "d0", "OR"
  )

priordata = list( npapers = npapers, a=a, b=b, d = d, c1=c1, f1 = f1,
                  c2 = c2, f2 = f2, y0 = metadata[,2], n0 = metadata[,1],
                  y1 = metadata[,4], n1 = metadata[,3]
                  )

inits = rep(list(list(
  beta   = rep(0,npapers),  
  delta   = rep(0,npapers),  
  alpha = 0, 
  d0 = 0,
  sigmainv2 = 1,
  tauinv2 = 1
)), 5)

run1 = jags(priordata, inits, parameters, "syncope1.txt", 
            n.chains=5, n.iter=1100, n.burnin=0, n.thin=1)
#proc.time()
names(run1)
Output1=AddBurnin(run1$BUGSoutput$sims.array,burnin=100,n.thin=1)

names(Output1)
print(Output1$Burnin.Summary)
colnames(Output1$Burnin.sims.matrix)

```

##### Problem 0. 
- Repeat everything with the reparameterized model. Now how is the convergence? 
```{r}
# Here is a reparameterized model. 
#Meta-analysis model
# sink("syncope1_repara.txt")
# cat("
#     model
#     {
#     for( i in 1:npapers ) {
#     y0[i] ~ dbin(pie0[i],n0[i])
#     y1[i] ~ dbin(pie1[i],n1[i])
#     logit(pie0[i]) = beta[i] - delta[i]/2
#     logit(pie1[i]) = beta[i] + delta[i]/2
#     beta[i]  ~ dnorm(alpha , sigmainv2)
#     delta[i] ~ dnorm(d0, tauinv2  )
#     OR[i]    = (pie1[i]/(1 - pie1[i])) / (pie0[i]/(1 - pie0[i]))
#     }
#     alpha ~ dnorm(a, b)
#     d0    ~ dnorm(0, d)
#     sigmainv2 ~ dgamma(c1,c2)
#     tauinv2   ~ dgamma(f1,f2)
#     sigma = 1/sqrt(sigmainv2)
#     tau   = 1/sqrt(tauinv2) 
#     }
#     ",fill = TRUE)
# sink()
```

```{r}
#repeat everything
run2 = jags(priordata, inits, parameters, "syncope1_repara.txt", 
            n.chains=5, n.iter=1100, n.burnin=0, n.thin=1)
names(run2)
Output2=AddBurnin(run2$BUGSoutput$sims.array,burnin=100,n.thin=1)

names(Output2)
print(Output2$Burnin.Summary)
colnames(Output2$Burnin.sims.matrix)

maindata = Output2$Burnin.sims.matrix[,c(10,11,31,32)]
dim(maindata)  #5000,4
head(maindata)


plot(1:5000, maindata[,1], ylab = "alpha", type="l")
plot(1:5000, maindata[,2], ylab = "d0"   , type="l")
plot(1:5000, maindata[,3], ylab = "sigma", type="l")
plot(1:5000, maindata[,4], ylab = "tau"  , type="l")
```

- According to the new plots, the convergence became better, especially alpha. 

#### Problem 1.
##### 1a. What is the change in the model from first version to second version? 

- The second version, we use $ \beta_i^* \sim N(\alpha,\sigma^2)$.

##### 1b. Is the model any different? 

- Same model. Different parametrization, prior distribution for new betas also changed.

##### 1c. Have any parameters changed meaning? Which are they, what are the changes?

- Yes. $\beta $ has been changed. 
- In the first model, $\beta_i $ is a study level random effect modeling differences between studies, with mean of 0.
- In the second model, $\beta_i^* $ has mean of $\alpha$, which is an overall success parameter. 

##### 1d. Is the posterior of any of the parameters any different from the original model? 

- The posterior of betas changed, since the prior distributions for betas were different.

##### 1e. What are differences between the output of the first model and the second version? 

- Second model showed better convergence. In addition, the $\alpha$ changed from -1.69 (model 1) to -2.41 (model 2).

#### Problem 2.

##### 2a. What is the single parameter are we most interested in? Has it changed meaning between the two model versions? Use model 2 for the remainder of this problem. 

- The key parameter that we are interested in is $\delta_0$. $\delta_0$ is the average, across papers, on the logit scale, of the log odds of the treatment (covariate) effect.
- The meaning of $\delta_0$ has not been changed.

##### 2b. What is your conclusion? Report your conclusion as an Odds Ratio and 95% interval. Do people with prior Heart Disease have better,  the same or worse outcomes after visiting the emergency room for syncope compared to those without heart disease? Give both a quantitative and a qualitative answer. 

- The OR mean is 2.68. The 95% interval is (1.59, 4.20). Prior of $\delta_0$ is centered at zero. ORs greater than 1 indicate greater odds of a bad outcome, so people with prior Heart Disease have worse outcomes after visiting the emergency room for syncope compared to those without heart disease.

#### Problem 3. 
##### Use the 2nd model for this problem. As sensitivity analysis for a meta-analysis, we rerun the analysis omitting each paper in turn. If there are $n$ papers contributing data to the analysis, we run $n$ additional analyses. A paper, that when omitted, changes our conclusions substantially is considered an influential paper. 
```{r}
n <- 8
parameters= c(
  "pie0", "pie1", "alpha", "sigma", 
  "tau", "d0", "OR"
  )
inits = rep(list(list(
        beta   = rep(0, n),  
        delta   = rep(0, n),  
        alpha = 0, 
        d0 = 0,
        sigmainv2 = 1,
        tauinv2 = 1
)), 5)

d_0 <- matrix(0,nrow = 9,ncol = 5)

for (i in 1:9) {
        metadata1 <- metadata[-i,]
        priordata1 = list(npapers = n, a=a, b=b, d = d, c1=c1, f1 = f1,
                             c2 = c2, f2 = f2, y0 = metadata1[,2], n0 = metadata1[,1],
                             y1 = metadata1[,4], n1 = metadata1[,3]
        )
        run3 = jags(priordata1, inits, parameters, "syncope1_repara.txt", 
                    n.chains=5, n.iter=1100, n.burnin=0, n.thin=1)

        Output3=AddBurnin(run3$BUGSoutput$sims.array,burnin=100,n.thin=1)
        d01 = Output3$Burnin.sims.matrix[,10]
        d_0[i,] = round(mysummary(d01), 3)
}
colnames(d_0) = c("mean", "sd", "2.5%", "97.5%", "p")
rownames(d_0) = c("paper1", "paper2", "paper3", "paper4", "paper5", "paper6", "paper7", "paper8", "paper9")

```

##### 3a. Report a table of the most important inference and how it changes as we delete each paper in turn. Clearly label and format your table. 
```{r}
kable(d_0)
```


##### 3b. Which paper is most influential? 

* The paper 9 is most influential

##### 3c. What is your conclusion? Is the final inference sensitive to omitting individual papers? 

* Not very sensitive.






