---
title: "LAB_3"
output:
  word_document: default
  html_document: default
---
#### Name Huiyu Hu
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
# Set working directory
setwd("/Users/huiyuhu/Desktop/Study/UCLA_Biostat/BIOSTAT234/lab/Lab 3")
getwd()

#LOAD NECESSARY PACKAGES
library(R2jags)
library(lattice)
#Function to help with the burn in 
load("AddBurnin.RData")
TraumaData = read.table("bcj97data.txt")

#Give the columns useful names 
colnames(TraumaData) <- c("death", "n",	"intercept",	"iss",	"rts", "age", "ti", "age*ti")

#first 6 observations from proto-typical cases 
head(TraumaData, n=6)
#For the 6 proto-typical cases define Xp the design matrix, Yp the outcomes (death=1), and np the number of trials
Xp = as.matrix(TraumaData[1:6,3:8])
Yp = TraumaData[1:6,1]
np = TraumaData[1:6,2]
#define the inverse of the Xp matrix to be used to convert the prior distributions on pi to distributions on the regression coefficients
invXp = solve(Xp)
#For the observed data define the design matrix, outcomes, and number of trials
Xobs = as.matrix(TraumaData[7:306,3:8])
Yobs = TraumaData[7:306,1]
nobs = TraumaData[7:306,2]
```
* Store the model in the file LAB3.Priors.txt
```{r}
# sink("LAB3.Priors.txt")
# cat("
# model{
# 	
# 	betas<-invXp %*% logitp[]
# 
# 	for(j in 1:6){
# 		logitp[j]<-logit(pie[j])
# 	}
# 	pie[1]~dbeta(1.1,8.5)
# 	pie[2]~dbeta(3.0,11.0)
# 	pie[3]~dbeta(5.9,1.7)
# 	pie[4]~dbeta(1.3,12.9)
# 	pie[5]~dbeta(1.1,4.9)
# 	pie[6]~dbeta(1.5,5.5)
# 
# }
# 
#   ",fill = TRUE)
# sink()
```

```{r}
#Now we incorporate the data into the model

# sink("Lab3.Posteriors.txt")
# cat("
# model{
# 	
# 	betas<-invXp %*% logitp[]
# 
# 	for(j in 1:6){
# 		logitp[j]<-logit(pie[j])
# 	}
# 	pie[1]~dbeta(1.1,8.5)
# 	pie[2]~dbeta(3.0,11.0)
# 	pie[3]~dbeta(5.9,1.7)
# 	pie[4]~dbeta(1.3,12.9)
# 	pie[5]~dbeta(1.1,4.9)
# 	pie[6]~dbeta(1.5,5.5)
# 
# 	
# 		for(i in 1:T){
# 		y[i] ~ dbern(p[i])
# 		p[i]<-ilogit(inprod(x[i,],betas[]))
# 		}
# 			
# 	
# }
#   ",fill = TRUE)
# sink()
```
#### To-Do 1
```{r}
ex1.data=list(invXp=invXp)
ex1.inits=rep(list(list(pie=c(0.5,0.5,0.5,0.5,0.5,0.5))),5)
ex1.parameters = c("betas", "pie[1:6]")

#Run the JAGS model
hw1.out = jags(ex1.data, ex1.inits, ex1.parameters, "LAB3.Priors.txt", 
	n.chains=5, n.iter=51000, n.burnin=0, n.thin=2, DIC=F) # the first 1000 iterations as a burn in

#Treat the first 1000 iterations as a burn in	
Output.hw1 = AddBurnin(hw1.out$BUGSoutput$sims.array,burnin=1000,n.thin=2)

ex2.data = list(x=Xobs, y=Yobs, T=300, invXp=invXp)
ex2.inits = rep(list(list(pie=c(0.5,0.5,0.5,0.5,0.5,0.5))),5)
ex2.parameters = c("betas", "pie[1:6]")


hw2.out = jags(ex2.data, ex2.inits, ex2.parameters, "Lab3.Posteriors.txt", 
	n.chains=5, n.iter=51000, n.burnin=0, n.thin=2, DIC=F)

#Treat the first 1000 iterations as a burn in	
Output.hw2 = AddBurnin(hw2.out$BUGSoutput$sims.array,burnin=1000,n.thin=2)

# Partial data T = 100
ex3.data = list(x=Xobs, y=Yobs, T=100, invXp=invXp)
ex3.inits = rep(list(list(pie=c(0.5,0.5,0.5,0.5,0.5,0.5))),5)
ex3.parameters = c("betas", "pie[1:6]")


hw3.out = jags(ex2.data, ex2.inits, ex2.parameters, "Lab3.Posteriors.txt", 
	n.chains=5, n.iter=51000, n.burnin=0, n.thin=2, DIC=F)
#Treat the first 1000 iterations as a burn in	
Output.hw3 = AddBurnin(hw3.out$BUGSoutput$sims.array,burnin=1000,n.thin=2)

```

1.	At what lags do the autocorrelations hit zero for the 6 regression coefficients? Are the beta autocorrelations better or worse than the 6 pi’s? 
```{r}
par(mfrow=c(2,3))
acf(hw2.out$BUGSoutput$sims.array[1:5000,1,1], main="beta1", lag.max = 200)
acf(hw2.out$BUGSoutput$sims.array[1:5000,1,2], main="beta2", lag.max = 200)
acf(hw2.out$BUGSoutput$sims.array[1:5000,1,3], main="beta3", lag.max = 200)
acf(hw2.out$BUGSoutput$sims.array[1:5000,1,4], main="beta4", lag.max = 200)
acf(hw2.out$BUGSoutput$sims.array[1:5000,1,5], main="beta5", lag.max = 200)
acf(hw2.out$BUGSoutput$sims.array[1:5000,1,6], main="beta6", lag.max = 200)
title(outer = T, "Beta 1-6 Autocorrelation Plots")
```
 Autocorrelation of $\beta_1$ hits 0 at Lag $\approx 100$  
 Autocorrelation of $\beta_2$ hits 0 at Lag $\approx 120$  
 Autocorrelation of $\beta_3$ hits 0 at Lag $\approx 100$  
 Autocorrelation of $\beta_4$ hits 0 at Lag $\approx 100$  
 Autocorrelation of $\beta_5$ hits 0 at Lag $\approx 70$  
 Autocorrelation of $\beta_6$ hits 0 at Lag $\approx 100$  
 
```{r}
par(mfrow=c(2,3))
acf(hw2.out$BUGSoutput$sims.array[1:5000,1,7], main="pi1", lag.max = 300)
acf(hw2.out$BUGSoutput$sims.array[1:5000,1,8], main="pi2", lag.max = 300)
acf(hw2.out$BUGSoutput$sims.array[1:5000,1,9], main="pi3", lag.max = 300)
acf(hw2.out$BUGSoutput$sims.array[1:5000,1,10], main="pi4", lag.max = 300)
acf(hw2.out$BUGSoutput$sims.array[1:5000,1,11], main="pi5", lag.max = 300)
acf(hw2.out$BUGSoutput$sims.array[1:5000,1,12], main="pi6", lag.max = 300)
```
 Autocorrelation of $\pi_1$ hits 0 at Lag $\approx 80$  
 Autocorrelation of $\pi_2$ hits 0 at Lag $\approx 100$  
 Autocorrelation of $\pi_3$ hits 0 at Lag $\approx 200$  
 Autocorrelation of $\pi_4$ hits 0 at Lag $\approx 90$  
 Autocorrelation of $\pi_5$ hits 0 at Lag $\approx 100$  
 Autocorrelation of $\pi_6$ hits 0 at Lag $\approx 250$  

2.	Turn in your properly formatted table of output for the full data set, and turn in a set of the 6 plots of the prior and posterior for the betas.

* Table of output

```{r}
names <- paste0("betas[", 1:6, "]")
Parameter <- c("Intercept", "ISS", "RTS", "Age", "ti", "Age * ti")
table1 <- cbind(Parameter, round(Output.hw2$Burnin.Summary[names, ], 4))
knitr::kable(table1, caption = "Posterior of Betas from Full Dataset Run")
```

* A set of the 6 plots of the prior and posterior for the betas
```{r}
temp3 = Output.hw1$Burnin.sims.matrix
temp4 = Output.hw2$Burnin.sims.matrix
par(mfrow=c(2,3))
plot(density(temp4[,1]),main="",xlab="beta1")  
# beta1.  
lines(density(temp3[,1],bw=.055),lty=2,lwd=2, col="blue")  
# prior
legend("topleft", legend=c("post","prior"), col=c("black", "blue"), lty=1:2 , lwd=c(1,2))
plot(density(temp4[,2]),main="",xlab="beta2")  
# beta2.  
lines(density(temp3[,2],bw=.05),lty=2,lwd=2, col="blue")  
# prior
plot(density(temp4[,3]),main="",xlab="beta3")  
# beta3.  
lines(density(temp3[,3],bw=.075),lty=2,lwd=2, col="blue")  
# prior
plot(density(temp4[,4]),main="",xlab="beta4")  
# beta4.  
lines(density(temp3[,4],bw=.045),lty=2,lwd=2, col="blue")  
# prior
plot(density(temp4[,5]),main="",xlab="beta5")  
# beta5.  
lines(density(temp3[,5],bw=.085),lty=2,lwd=2, col="blue")  
# prior
plot(density(temp4[,6]),main="",xlab="beta6")  
# beta6.  
lines(density(temp3[,6],bw=.085),lty=2,lwd=2, col="blue")  
# prior
```

3.	Turn in the results of the TODO step 2 properly formatted and your figures nicely annotated. 
TODO STEP2:	Compare output for 3 different runs: (i) prior (no data), (ii) partial data (T=100) posterior and then (iii) with the full data (T=300). 
a.	Compare all estimates and standard deviations in a table. [Given that you have three runs to compare, you’ll not want to have as many summaries of the posteriors as if you were doing a table of a single model. Think about how to best arrange the numbers in the table.]
```{r}
names <- paste0("betas[", 1:6, "]")
col <- c('mu.vect', 'sd.vect')
prior <-  round(Output.hw1$Burnin.Summary[names, col], 4)
p.post <- round(Output.hw2$Burnin.Summary[names, col], 4)
f.post <- round(Output.hw3$Burnin.Summary[names, col], 4)
table2 <- rbind(prior, p.post, f.post)
rownames(table2) <- c()
parameter <- rep(c("Beta 1", "Beta 2", "Beta 3", "Beta 4", "Beta 5", "Beta 6"), 3)
table2 <- cbind(parameter, table2)
rownames(table2) = c("Prior", " ", " ", " ", " ", " ", "Partial Posterior", " ", " ", "", " ", " ", "Full Posterior", " ", " ", "", " ", " ")
knitr::kable(table2, caption = "Parameter estimates from Prior, Partial Data, and Full Data models")
```

b.  Draw plots with three densities for each coefficient. The plot should show the prior density, the partial data posterior and full data posteriors. Label appropriately

```{r}
temp3 = Output.hw1$Burnin.sims.matrix
temp4 = Output.hw2$Burnin.sims.matrix
temp5 = Output.hw3$Burnin.sims.matrix
par(mfrow=c(2,3))
plot(density(temp4[,1]),main="",xlab="beta1")  
# beta1.  
lines(density(temp3[,1],bw=.055),lty=2,lwd=2, col="blue")  
# prior
lines(density(temp5[,1],bw=.055),lty=2,lwd=3, col="red")  
legend("topleft", lty = c(1,3,2), col = c("black", "red", "blue"), lwd = c(2,2,2), legend = c("Full posterior", "Partial posterior", "Prior"), bty = "n")

plot(density(temp4[,2]),main="",xlab="beta2")  
# beta2.  
lines(density(temp3[,2],bw=.05),lty=2,lwd=2, col="blue")  
# prior
lines(density(temp5[,2],bw=.055),lty=2,lwd=3, col="red")  

plot(density(temp4[,3]),main="",xlab="beta3")  
# beta3.  
lines(density(temp3[,3],bw=.075),lty=2,lwd=2, col="blue")  
# prior
lines(density(temp5[,3],bw=.055),lty=2,lwd=3, col="red")  

plot(density(temp4[,4]),main="",xlab="beta4")  
# beta4.  
lines(density(temp3[,4],bw=.045),lty=2,lwd=2, col="blue")  
# prior
lines(density(temp5[,4],bw=.055),lty=2,lwd=3, col="red")  

plot(density(temp4[,5]),main="",xlab="beta5")  
# beta5.  
lines(density(temp3[,5],bw=.085),lty=2,lwd=2, col="blue")  
# prior
lines(density(temp5[,5],bw=.055),lty=2,lwd=3, col="red")  

plot(density(temp4[,6]),main="",xlab="beta6")  
# beta6.  
lines(density(temp3[,6],bw=.085),lty=2,lwd=2, col="blue")  
# prior
lines(density(temp5[,6],bw=.055),lty=2,lwd=3, col="red")

```


4.	Turn in your answer to TODO step 4:	The model tracks the parameters $\pi_1$ to $\pi_6$, what is the interpretation of these parameters once the data has been incorporated?

- $\pi_1$ to $\pi_6$ are used to estimate the survival probability that 6 patients or predict the survival probability the patients  with same condition as these 6 patients.

Extra credit: you may (but don’t need to) Turn in your answer to TODO step 3.(1)	Extra credit 1: add code to calculate the probability that someone with Xs of 
1	2	7.55	25	0	0
1	11	7.8408	42	1	42
1	16	7.8408	80	1	80
	will die. 
```{r}
# sink("Lab3.Posteriors.pred.txt")
# cat("
# model{
# 
# 	betas<-invXp %*% logitp[]
# 
# 	for(j in 1:6){
# 		logitp[j]<-logit(pie[j])
# 	}
# 	pie[1]~dbeta(1.1,8.5)
# 	pie[2]~dbeta(3.0,11.0)
# 	pie[3]~dbeta(5.9,1.7)
# 	pie[4]~dbeta(1.3,12.9)
# 	pie[5]~dbeta(1.1,4.9)
# 	pie[6]~dbeta(1.5,5.5)
# 
# 
# 		for(i in 1:T){
# 		y[i] ~ dbern(p[i])
# 		p[i]<-ilogit(inprod(x[i,],betas[]))
# 		}
# for(k in 1:3) {
#       predPi[k] = ilogit(inprod(newx[k,], betas[]))
#   }
# 
# }
#   ",fill = TRUE)
# sink()
```

```{r}
x1 <- c(1,2,7.55,25,0,0)
x2 <- c(1,11,7.8408,42,1,42)
x3 <- c(1,16,7.8408,80,1,80)
newx <- rbind(x1, x2, x3)
colnames(newx) = colnames(Xobs)
hw4.data = list(x=Xobs, y=Yobs, T = 300, invXp=invXp, newx = newx)
hw4.inits = rep(list(list(pie=c(0.5,0.5,0.5,0.5,0.5,0.5))),5)
hw4.parameters = c("betas", "pie[1:6]", "predPi[1:3]")
hw4.out2 = jags(hw4.data, hw4.inits, hw4.parameters, "Lab3.Posteriors.pred.txt", n.chains=5, n.iter=51000, n.burnin=0, n.thin=2, DIC=F)
Output.hw4 = AddBurnin(hw4.out2$BUGSoutput$sims.array,burnin=1000,n.thin=2)

predPi <- paste0("predPi[", 1:3, "]")
table1 <- cbind(Parameter, round(Output.hw4$Burnin.Summary[predPi, ], 4))
knitr::kable(table1, caption = "Posterior of prediction of Pi")


```



