---
title: "Biostat234_Lab 5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#LOAD NECESSARY PACKAGES
setwd("/Users/huiyuhu/Desktop/Study/UCLA_Biostat/BIOSTAT234/lab/Lab 5")
library(R2jags)
load("AddBurnin.RData")
library(lattice)
load("model1data.RData")
```

```{r}
# sink("model1.txt")
# cat("
# # Response is number of Exencephaly malformations 
# # out of the total number of live fetuses
# # in the litter.  
# # Model #1.  
# model
# {
#         for (i in 1:N)
#         { 
#                 y[i] ~ dbin(p[i], n[i])
#                 logit(p[i]) <-  alpha[1] +alpha[2]*(z[i]-695) + beta[i] + 
#                                 x1[i]*alpha[3] +x2[i]*alpha[4]
#                 beta[i] ~ dnorm(0, tau)
#         }
#         for (j in 1:4)
#         {
#                 alpha[j] ~ dnorm(0, 0.001)
#         }
#         sigsq ~ dunif(0,6)
#         tau <- 1/sigsq
#         sigma <- sqrt(sigsq)
# 	for(j in 1:npred)
# 		{
# 			lppred[j] <- alpha[1] + alpha[2]*(pred[j]-695) + betanew
# 			logit(probpred[j]) <- lppred[j]
# 		}
# 		betanew ~ dnorm(0,tau)
# }
#     ",fill = TRUE)
#sink()

# Initial Values
model1inits = rep(
  list(list(alpha=c(0, 0, 0, 0), sigsq=1,
beta=c(0,0,0,1,0, 0,0,0,-1,0, 0,0,0,0,0, 0,0,0,0,0,
             0,0,0,0,0, 0,0,0,1,0, 0,0,0,0,0, 0,0,0,-1,0,
             0,0,0,0,1, 0,0,0,0,0, 0,0,0,0,0, 0,0,-1,0,0,
             0,0), betanew=0 )), 3)
```
```{r}
# sink("model2.txt")
# cat("
# # Model #2  
# model
# {
#         for (i in 1:N)
#         { 
#                 y[i] ~ dbin(p[i], n[i])
#                 p[i] ~ dbeta(a[i],b)
#                 a[i] <- b * exp(alpha[1] +alpha[2]*(z[i]-695) + 
#                         x1[i]*alpha[3] +x2[i]*alpha[4])
#         }
#         for (j in 1:4)
#         {
#                 alpha[j] ~dnorm(0, 0.001)
#         }
#         b ~ dgamma(ba,bb)
# }
#     ",fill = TRUE)
# sink()

```
```{r}

model1parameters= c("alpha", "p", "probpred", "beta", "tau", "sigma")

#### Did you load the data yet???

run1=jags(model1data, model1inits, model1parameters, "model1.txt", n.chains=3, n.iter=11000, n.burnin=0, n.thin=1, DIC=F)

Output1=AddBurnin(run1$BUGSoutput$sims.array,burnin=1000,n.thin=1)
```

Lab 5 Homework

1.	Print model 1 and on the printout, identify the following:  
a.	Where the likelihood is modeled;

`logit(p[i]) <-  alpha[1] +alpha[2]*(z[i]-695) + beta[i] +  x1[i]*alpha[3] +x2[i]*alpha[4]`

b.	Where the priors are modeled;

`        for (j in 1:4)`
`        {`
`                alpha[j] ~ dnorm(0, 0.001)`
`        }`
`        sigsq ~ dunif(0,6)`
`        tau <- 1/sigsq`
`        sigma <- sqrt(sigsq)`

c.	Where the linear predictor is specified;

`lppred[j] <- alpha[1] + alpha[2]*(pred[j]-695) + betanew`

d.	Where the predictions are modeled;

`for(j in 1:npred)`
` 		{`
`			lppred[j] <- alpha[1] + alpha[2]*(pred[j]-695) + betanew`
` 			logit(probpred[j]) <- lppred[j]`
`		}`
` 		betanew ~ dnorm(0,tau)`

e.	Where what predictions we are interested in are given.  

`logit(probpred[j]) <- lppred[j]`

2.	Give time series plots and autocorrelation plots for sigma and for alpha[2].  Give an estimate of how many iterations are sufficient for good estimation.  

```{r}
par(mfrow=c(1,2))
acf(Output1$Burnin.sims.array[,1,2], main="", lag.max=200) 
mtext("alpha[2]",side=3, line=1, cex=.8)
acf(Output1$Burnin.sims.array[,1,137], main="", lag.max = 200) 
mtext("sigma", side=3, cex=.8, line=1)
```

- According to the plot, 100 iterations are sufficient for good estimation.

3.	Which p[i] is largest?  Discuss (1-2 sentences) how you decide this, as there is more than one way to determine this. 

- I use which.max function to find the maximum from output. p[25] is the largest and the boxplots can also showed that p[25] is the largest.

```{r}
names <- paste0("p[", 1:62, "]")
p <- Output1$Burnin.Summary[names, ]
which.max(p[,1])
max(p[,1])
```
```{r}
#Boxplots for p
par(mfrow=c(1,1)) # sets plots in a 1x1 grid
#You may need to change the window size of the plot
boxplot(Output1$Burnin.sims.matrix[,c(seq(67,128))], range=0, axes=F)
axis(side=1,at=seq(1,62,5), labels=paste("p",seq(1,62,5)))
axis(side=2, at=0:5/5, labels=0:5/5)
box()
```


4.	Which beta[i] is largest?  Does it correspond to the same case as in Question 2?  Explain: Do the largest beta[i] and the largest p[i] have to correspond to the same case [i]?  

- By using which.max function and boxplot, beta[29] is the largest. No, the largest beta[i] and the largest p[i] do not necessarily have to correspond to the same case [i]. 

```{r}
names <- paste0("beta[", 1:62, "]")
beta <- Output1$Burnin.Summary[names, ]
which.max(beta[,1])
max(beta[,1])
```
```{r}
#Boxplots for Beta ## should include in the hw2
boxplot(Output1$Burnin.sims.matrix[,c(seq(5,66))], range=0, axes=F)
axis(side=1,at=seq(1,62,5), labels=paste("Beta",seq(1,62,5)))
axis(side=2, at=c(-5,0,5), labels=c(-5,0,5))
box()
```


5.	Plot posterior estimates of (.05, median, .95) for each prediction against blood glucose level.  Connect the dots between neighboring glucose levels.  Label and document your plot carefully.  

```{r}
#Plot of the predictions at each maternal blood glucose level
boxplot(Output1$Burnin.sims.matrix[,129:136], range=0, axes=F)
axis(side=1,at=1:8, labels=3:10*100)
axis(side=2, at=0:5/5, labels=0:5/5)
title(main = 'posterior for prediction against blood glucose level')
box()

names <- paste0("probpred[", 1:8, "]")
glucose.level <- c(300,400,500,600,700,800,900,1000)
table1 <- cbind(glucose.level, round(Output1$Burnin.Summary[names, ], 4))
knitr::kable(table1, caption = "Posterior of for each prediction against blood glucose level")
```

```{r}
#Extract the 95% CI and median from our BUGS summary table
r1<-as.matrix(run1$BUGSoutput$summary)
r1 <-r1[129:136,c(3,5,7)]

# Restructure the data so it's easy to plot
r2 <- data.frame(seq(100,800,100),rep(c("2.5%","Median","97.5%"),each=8),c(r1[,1],r1[,2],r1[,3]))
names(r2) <- c("Predicted","Percentile","Value")
r2$Predicted <- as.factor(r2$Predicted)

# Use ggplot2 to make it fancy
library(ggplot2)
ggplot(data=r2,aes(x=Predicted,y=Value,colour=Percentile)) + geom_point(size = 4) +
  geom_line(position=position_dodge(.1),aes(group=Percentile)) +
  theme(panel.grid.minor=element_blank()) + 
  labs(title = "Plot of Percentiles for Predicted Probabilites", 
       x = "Glucose", y = "Probability") +
  scale_colour_discrete(breaks=c("97.5%","Median","2.5%")) + ylim(0,1.0)
```


Extra credit.
6.	Repeat questions 1-4 for model 2.  One question does not make sense.  Which is it?  Why doesn’t it make sense?  

* There is no probpred in model 2. 

```{r}
load("model2data.RData")

# Initial values.  
model2inits = rep(list(list(alpha=c(0,0,0,0),b=1, 
p=c(.5,.5,.5,.4,.5, .5,.5,.5,.5,.5,
.5,.5,.5,.5,.5,.5,.5,.6,.5,.5,
.5,.5,.5,.4,.5,.5,.5,.5,.5,.5,
.5,.5,.5,.6,.5,.5,.5,.5,.5,.5,
.5,.5,.5,.4,.5,.5,.5,.5,.5,.5,
.5,.5,.5,.6,.5,.5,.5,.5,.5,.5, .5,.5))), 5)

#Rerun the model but with more iterations and a larger thinning parameter 
run2=jags(model2data, model2inits, model1parameters, "model2.txt", n.chains=5, n.iter=201000, n.burnin=0, n.thin=20, DIC=F)

Output2=AddBurnin(run2$BUGSoutput$sims.array,burnin=1000,n.thin=20)
```
a. Where the likelihood is modeled;

`p[i] ~ dbeta(a[i],b)`

b. Where the priors are modeled;

` alpha[j] ~dnorm(0, 0.001)`
`b ~ dgamma(ba,bb)`

c. Where the linear predictor is specified;

`a[i] <- b * exp(alpha[1] +alpha[2]*(z[i]-695) + x1[i]*alpha[3] +x2[i]*alpha[4])`

d. Where the predictions are modeled; (There is no prediction part in the model)

e. Where what predictions we are interested in are given.
`p[i]`

Give time series plots and autocorrelation plots for sigma and for alpha[2]. Give an estimate of how many iterations are sufficient for good estimation.

- There is no sigma. But for alpha[2], 30 iterations are sufficient for good estimation.

```{r}
acf(Output2$Burnin.sims.array[,1,2], main="", lag.max=200) 
mtext("alpha[2]",side=3, line=1, cex=.8)

```

Which p[i] is largest? Discuss (1-2 sentences) how you decide this, as there is more than one way to determine this.

- p[38] is the largest.

```{r}
names <- paste0("p[", 1:62, "]")
p <- Output2$Burnin.Summary[names, ]
which.max(p[,1])
max(p[,1])
```

Which beta[i] is largest? 

- beta[29] is the largest.

```{r}
names <- paste0("beta[", 1:62, "]")
beta <- Output1$Burnin.Summary[names, ]
which.max(beta[,1])
max(beta[,1])
```



7.	The notes suggest possibly taking 1/(b+1) ~ uniform(0,1).  Run this variation of model 2.  Do this by defining c = 1/(b+1), then calculating b as a function 
of c.  Provide your model code, the two posteriors for b (preferably on same plot, but not required), and the posterior summary info for parameter b.  Indicate on your model code the parts 1a through 1e.  

